The Cluster Difference Imaging Photometric Survey (CDIPS). I. Light Curves \& Planet Candidates in Open Clusters from TESS Sectors 6 \& 7

### 2019-06-23 paper/ms.tex
add vetting report to appendix

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{sector6_cam[1]_ccd[1-2-3-4]cluster_field_star_positions.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf  Zoom of Sector 6, Camera 1 from Figure~\ref{fig:lcradec}.}
		\label{fig:lcradeczoom}
	}
\end{figure}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{cdf_T_mag.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Cumulative number of CDIPS light curves a function of TESS
    $T$-band magnitude.}  Light curves were made for the subset of
    stars shown in Figure~\ref{fig:cdips_targets} that were observed
    in Sectors 6 or 7.
		\label{fig:cdf_T_mag}
	}
\end{figure}

@@

A number of the light curves show residual variability, often of stellar 
origin.
This expected for a sample of young stars; the issue of finding transits in 
the face of large stellar rotation signals 
has been explored by both \citet{rizzuto_zeitV_2017} and also 
\citet{hippke_wotan_2019}.
The former adopted an approach that we have yet to explore, which is to 
pass a sliding window over the light curve that, within each step, performs a 
Bayesian model comparison between a spline and a spline-plus-notch model.  If 
the spline-plus-notch model is favored, the transits are preserved for 
subsequent discovery.
\citet{hippke_wotan_2019}, conversely, described a large number of possible 
detrending methods, and found that many performed more or less the same at 
recovering planetary transits for a sample of young stars with strong rotation 
signals.

Our approach in this work is essentially identical to one of the methods 
described by \citet{hippke_wotan_2019}.
First, we run a generalized Lomb-Scargle periodogram (CITE) on each TFA light 
curve.
If a peak is found with false alarm probability below $10^{-5}$, we consider 
the star ``variable'', and opt to 
detrend the light curve with robust penalized B-splines, which are splines 
with knot-length automatically determined via cross-validation 
\citep{eilers_flexible_1996}. The idea behind the cross-validation is that a 
larger  number of knots leads to smaller residuals on training data, but 
larger errors 
when tested on the entire dataset.
We used the \texttt{wotan} implementation, which is a wrapper to the 
\texttt{pyGAM} spline fitter, with $2\sigma$ clipping of outliers from the fit 
residuals at each iteration 
\citep{serven_pygam_2018_1476122,hippke_wotan_2019}.
The maximum number of spline knots is set to 50, which for each TESS sector 
($\approx 25\,{\rm days}$) is commensurate with a $\approx0.5\,{\rm day}$ 
window.

@@

This yielded a few thousand light curves.  About two-thirds already had been 
detrended with splines, and were not processed further.
For the remaining third we
performed TFA signal reconstruction using 
\texttt{vartools}.
In this process, the model lightcurve $A(i)$ in 
Equation~\ref{eq:tfa_to_minimize} is set to the phase-binned signal from the 
most powerful peak in the TLS spectrum, rather than being a constant function. 
This helps improve lower the scatter of the light curve at frequencies near 
the TLS peak.

@@

The classifications we used assess:
\begin{enumerate}
    \item {\it The source of photometric variability}. Designations include 
    tags for planet candidates, eclipsing binaries, instrumental variations, 
    stellar variations, and ``weirdos''.
    \item {\it Cluster membership status}. By default, all light curves were 
    made for stars with at least one literature claim of cluster membership.  
    We therefore include tags only for non-cluster members, and possible 
    non-cluster members. Typically the primary source of this information is 
    the Gaia-DR2 parallax.
    \item {\it Photometric blends}. Tags are created to highlight whether the 
    depth of the photometric signal shows a strong dependence on aperture 
    size, and also whether the in-transit minus the out-of-transit images 
    reveal that the source of variability is in fact far from the target star.
\end{enumerate}
The actual classification process was performed by LGB, JH, and JNW.
The \texttt{TagSpaces} software was used~--~an extremely helpful tool for the 
purposes of easily assigning labels to documents.

@@

In \S~\ref{subsec:ctois}, we described the process by which we made 
\texttt{.pdf} documents suitable for assessing which objects were interesting 
enough to merit further study.

This section briefly summarizes this document. Updated versions and their 
\texttt{README} files will live at this web-address: 
\url{mast.stsci.edu/CDIPS}.

\begin{itemize}
\item {\it Page 1: period-search summary}. Periodograms from TLS and 
phase-dispersion minimization 
\citep{hippke_TLS_2019,stellingwerf_period_1978}, as calculated 
with \texttt{astrobase.periodbase} are shown. The top three peaks from each 
method are shown in the second and third rows; the raw light curve is in the 
top-right. A small finder chart from DSS is shown (CITE).

\item {\it Page 2: light curve diagnostics}.
\end{itemize}

>>>>

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.47\textwidth]{sector6_cam[1]_ccd[1-2-3-4]cluster_field_star_positions.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf  Zoom of Sector 6, Camera 1 from Figure~\ref{fig:lcradec}.}
		The most populated cluster is XXX, primarily due to candidate
		members proposed by \citet{dias_proper_2014}.
		\label{fig:lcradeczoom}
	}
\end{figure}

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.45\textwidth]{cdf_T_mag.png}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Cumulative number of CDIPS light curves a function of TESS
    $T$-band magnitude.}  Light curves were made for the
    target stars (Figure~\ref{fig:cdips_targets}) that were observed
    in Sectors 6 or 7.
		\label{fig:cdf_T_mag}
	}
\end{figure}

@@

A number of the light curves show residual variability, often of stellar 
origin.
This is expected for a sample of young stars; the problem of finding transits in 
the face of large stellar rotation signals 
has been explored by both \citet{rizzuto_zeitV_2017} and also 
\citet{hippke_wotan_2019}.
The former adopted an approach that we have yet to explore, which is to 
pass a sliding window over the light curve that, within each step, performs a 
Bayesian model comparison between a spline and a spline-plus-notch model.  If 
the spline-plus-notch model is favored, the transits are preserved for 
subsequent discovery.
\citet{hippke_wotan_2019}, conversely, described a number of different
detrending methods, and found that most performed more or less the same at 
recovering planetary transits for a sample of young stars with strong rotation 
signals.

Our approach in this work is essentially identical to one of the methods 
described by \citet{hippke_wotan_2019}.
First, we run a generalized Lomb-Scargle periodogram on each TFA light 
curve \citep{lomb_1976,scargle_studies_1982,vanderplas_periodograms_2015}.
If a peak is found with false alarm probability below $10^{-5}$, we consider 
the star ``variable'', and opt to 
detrend the light curve with robust penalized B-splines, which are splines 
with knot-length automatically determined via cross-validation 
\citep{eilers_flexible_1996}. The idea behind the cross-validation is that a 
larger  number of knots leads to smaller residuals on training data, but 
larger errors 
when tested on the entire dataset.
We used the \texttt{wotan} implementation, which is a wrapper to the 
\texttt{pyGAM} spline fitter, with $2\sigma$ clipping of outliers from the fit 
residuals at each iteration 
\citep{serven_pygam_2018_1476122,hippke_wotan_2019}.
The maximum number of spline knots is set to 50, which for each TESS sector 
($\approx 25\,{\rm days}$) is commensurate with a $\approx0.5\,{\rm day}$ 
window.

We verified by injecting and recovering a 2.5 mmag central-transit planet
with variable period that this detrending approach marginally improved
the recovery probability, compared to not doing anything.
The main cases for which it helps are those with large residual stellar
variability in the TFA light curve.

@@

This yielded a few thousand light curves.  About two-thirds already had been 
detrended with splines, and were not processed further.
For the remaining third we
performed TFA signal reconstruction using 
\texttt{vartools}.
In this process, the model lightcurve $A(i)$ in 
Equation~\ref{eq:tfa_to_minimize} is set to the phase-binned signal from the 
most powerful peak in the TLS spectrum, rather than being a constant function. 
This helps lower the scatter of the light curve at frequencies away from
the TLS peak.

@@

The classifications we used assessed:
\begin{enumerate}
    \item {\it Photometric variability classification}. Designations include 
    tags for planet candidates, eclipsing binaries, instrumental variations, 
    stellar variations, and ``weirdos''.
    %
    \item {\it Cluster membership status}. By default, all light curves were 
    made for stars with at least one literature claim of cluster membership.  
    However, some of these claims pre-dated the availability of Gaia-DR2
    parallaxes or proper motions.
    We therefore re-assess the membership status on a case-by-case basis 
    through inspection of the available astrometry and photometry.
    Known asterisms are also omitted at this point
    \citep[{\it e.g.},][]{sulentic_revised_1973,kos_galah_2018}.
    %
    \item {\it Photometric blends}. Tags are created to highlight whether the 
    depth of the photometric signal shows a strong dependence on aperture 
    size, and also whether the in-transit minus the out-of-transit images 
    reveal that the source of variability is in fact far from the target star.
\end{enumerate}
The actual classification process was performed by LGB, JH, and JNW.
The \texttt{TagSpaces} software was used~--~an extremely helpful tool for the 
purposes of easily assigning labels to documents.

@@

In \S~\ref{subsec:ctois}, we described the process by which we made 
vetting reports suitable for assessing which objects were interesting 
enough to merit further study.

Figures~\ref{fig:pg1} to~\ref{fig:pg6} of this
section summarize this document. Updated versions and their 
\texttt{README} files will live at this web-address: 
\url{mast.stsci.edu/CDIPS}.
The planet candidate chosen for these figures
(Gaia DR2 \texttt{5541111035713815552} = TIC \texttt{134528212}) was
chosen for the demonstration in part because it passes all
the tests.

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page01.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 1: period-search summary.} Periodograms from TLS and 
		phase-dispersion minimization, as calculated 
		with \texttt{astrobase.periodbase}, are shown in the top left and top center 
		\citep{bhatti_astrobase_2018,hippke_TLS_2019,stellingwerf_period_1978}.
		The top three peaks from each 
		method are shown in the second and third rows; the raw light curve is in the 
		top-right. A small finder chart from DSS is inset to the top left, with
		the 1.5-pixel radius aperture used to extract the light curve in orange (CITE).
		\label{fig:pg1}
	}
\end{figure*}

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page02.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 2: light curve diagnostics.} 
		Time-series of raw flux (\texttt{IRM2}),  TFA-detrended flux (\texttt{TF2}), stellar-variability
		detrended flux, and the background are shown as a function of
		barycentric Julian date.
		The overplotted dashed vertical lines are the ephemeris of the highest-power
		TLS peak from Figure~\ref{fig:pg1}.
		An important check is whether the flux dips are correlated with changes in the 
		background level -- in this case, they are not.
		The standard deviation and TESS magnitude are quoted in the upper right.
		The red line in the second from the top plot is the windowed 
		spline described in Section~\ref{subsec:ctois} -- in this case it is near its maximal
		frequency.
		\label{fig:pg2}
	}
\end{figure*}

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page03.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 3: Transit diagnostics.} The plots show the maximally-detrended
		light curve (top); the phase-folded light curve centered on the primary transit
		(middle left); the phase-folded light curve centered on the secondary eclipse
		(middle right); the odd-numbered transits (lower left); and the
		even-numbered transits (lower right).
		The stellar parameters ($T_{\rm eff}, R_\star$) are taken from Gaia-DR2 
		when available.
		The first eight lines of text are parameters determined 
		from the best-fitting TLS model.
		The one exception is the planet radius, which uses the stellar radius
		as noted above.
		The ``flux contamination'' (\texttt{TICCONT}) from neighboring stars
		is {\it never} taken into account, because transit depth dilution does
		not affect image subtraction analyses in the same manner as 
		aperture-photometry reductions.
		To estimate the transit to occulation
		depth ratio,
		the phase-folded light curve is also coarsely fit via maximum likelihood
		by a sum of two gaussians.
		``AstExc'' refers to the Gaia-DR2 astrometric excess, which can indicate
		hints of astrometric binarity in the system.
		``$R_\star + M_\star \rightarrow T_{b0}$'' gives the duration of a 
		zero-eccentricity central transit based on the Gaia-DR2 stellar radius,
		and a stellar mass interpolated from the \citet{pecaut_intrinsic_2013} table,
		under the assumption that the star is a dwarf.%\footnote{\url{http://www.pas.rochester.edu/~emamajek/EEM_dwarf_UBVIJHK_colors_Teff.txt}}.
		\label{fig:pg3}
	}
\end{figure*}

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page04.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 4: Light curves for increasing aperture sizes.} 
		Apertures of radius 1, 1.5, and 2.25 pixels are shown from top to bottom.
		The blue line is the reference transit depth from the best-fitting TLS model.
		Any changes in depth with aperture size might indicate the need for a 
		detailed pixel-level analysis.
		\label{fig:pg4}
	}
\end{figure*}

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page05.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 5: Cluster membership assessment diagnostics.}
		The star was considered a candidate cluster member by the source(s) listed under
		``Reference'', in this case \citet{dias_proper_2014}.
		The name used in the Dias catalog in this case was \texttt{264-025831}, a
		UCAC-4 identifier, which can be back-referenced to find that \citet{dias_proper_2014}
		assigned this star a
		membership probability in NGC 2546 of 98\%.
		The base catalog for the plots is chiefly that of \citet{Kharchenko_et_al_2013},
		due to its homogeneous parameter determination procedure (particularly for age).
		If a match to the \citet{Kharchenko_et_al_2013} catalog is found, then the remaining
		plots are populated.
		Top-left shows the parallax, with orange points sampled from the Gaia-DR2
		posterior, black points the other cluster members in the Kharchenko catalog,
		and the blue line the claimed Kharchenko parallax for the cluster.
		A large number of field contaminants are visible in this case.
		Top-right are the Gaia proper motions, where against black points are cluster
		members from Kharchenko, and the orange is the target star.
		Bottom-left is the HR diagram, and bottom-right are the on-sky positions.
		In the text,
		\texttt{N1sr2} is the number of $1\sigma$ cluster members claimed
		by \citet{Kharchenko_et_al_2013} within the cluster angular radius;
		$\log t$ is the base-10 logarithm of the age in years;
		\texttt{type} matches the type codes provided by \citet{Kharchenko_et_al_2013}.
		\label{fig:pg5}
	}
\end{figure*}

\begin{figure*}[!h]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.98\textwidth]{gaiatwo0005541111035713815552-0007_page06.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf Page 6: Imaging variability diagnostics.} 
		This page is intended to help diagnose which stars are producing the observed
		variability.
		Top-left and top-center are the mean out-of-transit (OOT) and mean in-transit 
		calibrated images (separate from any of our image-subtraction analysis).
		The OOT images are based on the same number of exposures as the in-transit images
		and split evenly before and after each transit \citep[following][]{bryson_identification_2013,kostov_l9859_2019}.
		The yellow star is the target; cyan dots are the flux-weighted centroid of the entire
		image for each transit event; small red crosses are WCS-projected locations of neighbor stars.
		Middle-left is the most important sub-panel: the difference between
		the OOT and in-transit mean images.
		If the variability shown in background map (units: ADU) is off-target, the
		transit is typically not from the target star.
		Middle-center is the same, normalized by the uncertainty map.
		Lower left and lower center show the DSS field in linear and log scales
		at roughly the same pixel scale as the TESS image,
		with the 1, 1.5, and 2.25 pixel-radius apertures in blue, orange, and green respectively.
		The brightness of neighborhood stars is given on the far right.
		Note the slight rotation difference between DSS and TESS images;
		DSS images are aligned north-up, east-left; TESS images are oriented as closely as
		possible to this system without actually performing the rotation.
		\label{fig:pg6}
	}
\end{figure*}
